---
title: "p8105_hw2_maf2311"
author: "Mackenzie Frost"
date: "Octobr 4 2022"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

```{r q1, include = FALSE}
nyc_subway = read_csv(file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, starts_with("route"), entry,
    exit_only, vending, entrance_type, ada
  ) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

NYC Transit Entry/Exit Dataset Description: 

The NYC Transit Entry/Exit Dataset has the following variables: `r names(nyc_subway)`. To clean the data, I first imported the csv file, made the names more uniform with the janitor package, and only kept the variables of interest. The binary variable 'entry' was changed to a logical variable using the mutate function. The NYC Transit Entry/Exit Dataset dataset has `r nrow(nyc_subway)` rows and `r ncol(nyc_subway)` columns. These data are tidier than when I first imported the dataset, however we can update the variable names (specifically the route variables) to make them more user friendly. 

```{r}
stn = 
  nyc_subway %>% 
  select(line, station_name) %>% 
  distinct

compliant = 
  nyc_subway %>% 
  filter(ada == TRUE) %>%
  select(line, station_name) %>% 
  distinct

ent=
  nyc_subway %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

* There are `r nrow(stn)` distinct stations. 
* There are `r nrow(compliant)` stations that are ADA compliant. 
* The proportion f station entrances/exits without vending that allow entrance is `r ent`.

# Problem 2

```{r}
mrTrashWheel = 
  read_excel("./Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", skip = 1, n_max = 547) %>%
  janitor::clean_names() %>%
  select(-x15, -x16) %>%
  mutate(sports_balls = round(sports_balls, digits = 0)) %>%
  mutate(sports_balls = as.integer(sports_balls)) %>%
  mutate(mrORprof = 1)

profTrashWheel = 
  read_excel("./Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", skip = 1, n_max = 94) %>%
  janitor::clean_names() %>%
  mutate(mrORprof = 2)

trashwheel = 
  merge(mrTrashWheel, profTrashWheel, all = TRUE)
```

The Trashwheel Dataset has the following variables: `r names(trashwheel)`. There are `r nrow(trashwheel)` observations. The variable 'mrORprof' distinguishes between Mr. Trash Wheel (1) and Professor Trash Wheel (2).

```{r}
filtered2 = 
  trashwheel %>%
  filter(mrORprof == 2)
weight = sum(filtered2$weight_tons)

filtered1 = 
  trashwheel %>%
  filter(mrORprof == 1 & year == "2020")
balls = sum(filtered1$sports_balls)
```

The total weight of trash collected by Professor Trash Wheel: `r weight` tons

The total number of sports balls collected by Mr. Trash Wheel in 2020: `r balls` sports balls

# Problem 3







