---
title: "p8105_hw2_maf2311"
author: "Mackenzie Frost"
date: "Octobr 4 2022"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

```{r q1, include = FALSE}
nyc_subway = read_csv(file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, starts_with("route"), entry,
    exit_only, vending, entrance_type, ada
  ) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

NYC Transit Entry/Exit Dataset Description: 

The NYC Transit Entry/Exit Dataset has the following variables: `r names(nyc_subway)`. To clean the data, I first imported the csv file, made the names more uniform with the janitor package, and only kept the variables of interest. The binary variable 'entry' was changed to a logical variable using the mutate function. The NYC Transit Entry/Exit Dataset dataset has `r nrow(nyc_subway)` rows and `r ncol(nyc_subway)` columns. These data are tidier than when I first imported the dataset, however we can update the variable names (specifically the route variables) to make them more user friendly. 

```{r}
stn = 
  nyc_subway %>% 
  select(line, station_name) %>% 
  distinct

compliant = 
  nyc_subway %>% 
  filter(ada == TRUE) %>%
  select(line, station_name) %>% 
  distinct

ent=
  nyc_subway %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

* There are `r nrow(stn)` distinct stations. 
* There are `r nrow(compliant)` stations that are ADA compliant. 
* The proportion f station entrances/exits without vending that allow entrance is `r ent`.

# Problem 2

```{r}
mrTrashWheel = 
  read_excel("./Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", skip = 1, n_max = 547) %>%
  janitor::clean_names() %>%
  select(-x15, -x16) %>%
  mutate(sports_balls = round(sports_balls, digits = 0)) %>%
  mutate(sports_balls = as.integer(sports_balls)) %>%
  mutate(mrORprof = 1)

profTrashWheel = 
  read_excel("./Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", skip = 1, n_max = 94) %>%
  janitor::clean_names() %>%
  mutate(mrORprof = 2)

trashwheel = 
  merge(mrTrashWheel, profTrashWheel, all = TRUE)
```

The Trashwheel Dataset has the following variables: `r names(trashwheel)`. There are `r nrow(trashwheel)` observations. The variable 'mrORprof' distinguishes between Mr. Trash Wheel (1) and Professor Trash Wheel (2).

```{r}
filtered2 = 
  trashwheel %>%
  filter(mrORprof == 2)
weight = sum(filtered2$weight_tons)

filtered1 = 
  trashwheel %>%
  filter(mrORprof == 1 & year == "2020")
balls = sum(filtered1$sports_balls)
```

The total weight of trash collected by Professor Trash Wheel: `r weight` tons

The total number of sports balls collected by Mr. Trash Wheel in 2020: `r balls` sports balls

# Problem 3

```{r}
pols_month = 
  read_csv(file = "./fivethirtyeight_datasets/pols-month.csv") %>%
  separate(col=mon, into=c('year', 'month', 'day'), sep='-') %>%
  mutate(month = month.abb[as.numeric(month)]) %>%
  mutate(president = ifelse(prez_gop == 0, "dem", "gop")) %>%
  select(-day, -prez_gop, -prez_dem)

snp = 
  read_csv(file = "./fivethirtyeight_datasets/snp.csv") %>%
  separate(col=date, into=c('month', 'day', 'year'), sep='/') %>%
  mutate(month = month.abb[as.numeric(month)]) %>%
  mutate(year = ifelse(year<23, paste("20",year, sep = ""), paste("19",year, sep = ""))) %>%
  select(year, month, close, -day)

unemp = 
  read_csv(file = "./fivethirtyeight_datasets/unemployment.csv") %>%
  rename(year = Year) %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "percent")

result = 
  full_join(pols_month, snp, by = c("year", "month")) %>%
  mutate(year = as.numeric(year))

final = full_join(result, unemp, by = c("year", "month"))
```

### Pols_month Dataset

This dataset contained information on the political breakdown of the United States Government from `r min(pols_month$year)` to `r max(pols_month$year)`. It had `r nrow(pols_month)` rows and `r ncol(pols_month)` columns, and the variables included:  `r names(pols_month)`

### SNP Dataset

This dataset contained information on monthly closing values of the Standard & Poor’s stock market index (S&P) from `r min(snp$year)` to `r max(snp$year)`. It had `r nrow(snp)` rows and `r ncol(snp)` columns, and the variables included:  `r names(snp)`

### Unemployment Dataset

This dataset contained information on monthly unemployment rates (as percentages) from `r min(unemp$year)` to `r max(unemp$year)`. It had `r nrow(unemp)` rows and `r ncol(unemp)` columns, and the variables included:  `r names(unemp)`

### Final Dataset

This dataset contains all information from the 3 datasets mentioned above, from `r min(final$year)` to `r max(final$year)`. The data entries are joined by the 'year' and 'month' variables, and there are `r ncol(final)` columns and `r nrow(final)` rows.







